---
title: "Modélisation mémoire Education"
author: "Davidson Adrien"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 4
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(scipen = 999)
```


# **Importing required packages**

```{r import packages}
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(glmulti))
suppressPackageStartupMessages(library(flextable))
suppressPackageStartupMessages(library(performance))
suppressPackageStartupMessages(library(Rcpp))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(see))
suppressPackageStartupMessages(library(skimr))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(dlookr))
suppressPackageStartupMessages(library(report))
suppressPackageStartupMessages(library(ggstatsplot))
suppressPackageStartupMessages(library(gt))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(gtsummary))
suppressPackageStartupMessages(library(ggfortify))

```


# **Importing the dataset for LR**

```{r import dataset}
df_mem <- read_excel('./Data_model2.xlsx')
data_contraint <- read_excel('./Occurence_contraintes.xlsx') 

```


# **Data wrangling**

## Factorizing the variables

```{r data wrangling: as.factor-ing the variables}
df_mem$depot <- as.factor(df_mem$depot)
df_mem$recoded_temps_depot <- as.factor(df_mem$recoded_temps_depot)
df_mem$sexe <- as.factor(df_mem$sexe)
df_mem$loc_etud_sec <- as.factor(df_mem$loc_etud_sec)
df_mem$type_etab_term <- as.factor(df_mem$type_etab_term)
df_mem$compos_seuh <- as.factor(df_mem$compos_seuh)
df_mem$domaine_etud1 <- as.factor(df_mem$domaine_etud1)
df_mem$domaine_etud2 <- as.factor(df_mem$domaine_etud2)
df_mem$generation <- as.factor(df_mem$generation)
df_mem$autr_etud <- as.factor(df_mem$autr_etud)
df_mem$approche <- as.factor(df_mem$approche)
df_mem$qual_accomp <- as.factor(df_mem$qual_accomp)
df_mem$temps_emploi <- as.factor(df_mem$temps_emploi)
df_mem$contrainte <- as.factor(df_mem$contrainte)

df_mem$contrainte_interet <- as.integer(df_mem$contrainte_interet)
df_mem$contrainte_interet <- as.factor(df_mem$contrainte_interet)

df_mem$contrainte_motivation <- as.integer(df_mem$contrainte_motivation)
df_mem$contrainte_motivation <- as.factor(df_mem$contrainte_motivation)

df_mem$contrainte_ordi <- as.integer(df_mem$contrainte_ordi)
df_mem$contrainte_ordi <- as.factor(df_mem$contrainte_ordi)

df_mem$contrainte_elect <- as.integer(df_mem$contrainte_elect)
df_mem$contrainte_elect <- as.factor(df_mem$contrainte_elect)

df_mem$contrainte_internet <- as.integer(df_mem$contrainte_internet)
df_mem$contrainte_internet <- as.factor(df_mem$contrainte_internet)

df_mem$contrainte_boulot <- as.integer(df_mem$contrainte_boulot)
df_mem$contrainte_boulot <- as.factor(df_mem$contrainte_boulot)

df_mem$contrainte_fam <- as.integer(df_mem$contrainte_fam)
df_mem$contrainte_fam <- as.factor(df_mem$contrainte_fam)

df_mem$contrainte_sante <- as.integer(df_mem$contrainte_sante)
df_mem$contrainte_sante <- as.factor(df_mem$contrainte_sante)

df_mem$contrainte_note <- as.integer(df_mem$contrainte_note)
df_mem$contrainte_note <- as.factor(df_mem$contrainte_note)

df_mem$contrainte_finance <- as.integer(df_mem$contrainte_finance)
df_mem$contrainte_finance <- as.factor(df_mem$contrainte_finance)

df_mem$contrainte_autr_etud <- as.integer(df_mem$contrainte_autr_etud)
df_mem$contrainte_autr_etud <- as.factor(df_mem$contrainte_autr_etud)

df_mem$contrainte_sujet_recherche <- as.integer(df_mem$contrainte_sujet_recherche)
df_mem$contrainte_sujet_recherche <- as.factor(df_mem$contrainte_sujet_recherche)

df_mem$contrainte_sujet_complexe <- as.integer(df_mem$contrainte_sujet_complexe)
df_mem$contrainte_sujet_complexe <- as.factor(df_mem$contrainte_sujet_complexe)

df_mem$contrainte_donnees <- as.integer(df_mem$contrainte_donnees)
df_mem$contrainte_donnees <- as.factor(df_mem$contrainte_donnees)

df_mem$contrainte_analyse_donnees <- as.integer(df_mem$contrainte_analyse_donnees)
df_mem$contrainte_analyse_donnees <- as.factor(df_mem$contrainte_analyse_donnees)

df_mem$contrainte_encadrement <- as.integer(df_mem$contrainte_encadrement)
df_mem$contrainte_encadrement <- as.factor(df_mem$contrainte_encadrement)

df_mem$contrainte_emigration <- as.integer(df_mem$contrainte_emigration)
df_mem$contrainte_emigration <- as.factor(df_mem$contrainte_emigration)

df_mem$contrainte_autre <- as.integer(df_mem$contrainte_autre)
df_mem$contrainte_autre <- as.factor(df_mem$contrainte_autre)

df_mem$info_tfe <- as.factor(df_mem$info_tfe)
df_mem$difficulte_tfe <- as.factor(df_mem$difficulte_tfe)
df_mem$utilite_tfe <- as.factor(df_mem$utilite_tfe)
df_mem$projet_personnel <- as.factor(df_mem$projet_personnel)
df_mem$choix_etud <- as.factor(df_mem$choix_etud)
df_mem$education_fam <- as.factor(df_mem$education_fam)

df_mem$consommation_pps <- as.integer(df_mem$consommation_pps)
df_mem$consommation_pps <- as.factor(df_mem$consommation_pps)

```


## Partitions of the dataset for modelling purposes

```{r Partitions of the dataset for modelling purposes}
#Partitionning the dataset
var1 <- c('sexe', 'loc_etud_sec', 'type_etab_term', 'compos_seuh', 'domaine_etud1',
          'autr_etud', 'generation', 'approche', 'qual_accomp', 'temps_emploi',  
          'info_tfe', 'difficulte_tfe', 'utilite_tfe', 'projet_personnel',
          'choix_etud', 'consommation_pps', 'education_fam', 'contrainte_motivation')

df1 <- df_mem %>%
  select(depot, var1)

df2 <- df_mem %>%
    select(colnames(df_mem), -var1, -c('soutenance', 'temps_soutenance',
                                     'duree_etud_pot', 'duree_etud_reel',
                                     'duree_etud_diff', 'consent', 'preciser_etab_term',
                                     'etud_1', 'tranche_age', 'fac',
                                     'etud_2', 'comment_accomp',
                                     'preciser_autr_contr', 'domaine_etud2',
                                     'comment_difficulte', 'comment_utilite',
                                     'promotion', 'recoded_temps_depot', 'contrainte')) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0))

df3 <- df_mem %>%
  filter(contrainte == "Oui") %>%
    select(colnames(df_mem),-var1, contrainte, -c('soutenance', 'temps_soutenance',
                                     'duree_etud_pot', 'duree_etud_reel',
                                     'duree_etud_diff', 'consent', 'preciser_etab_term', 'etud_1', 'tranche_age', 'fac', 'etud_2', 'comment_accomp',
                                     'preciser_autr_contr', 'domaine_etud2',
                                     'comment_difficulte', 'comment_utilite',
                                     'promotion', 'recoded_temps_depot', 'contrainte')) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0)) 

df1_col <- list(colnames(df1))
df2_col <- list(colnames(df2))

#variables in df1
df1_col

#Variables in df2
df2_col
```

# **Multivariate Analysis I**

## Logistic Regression

### Stepwise selection: logistic regression

```{r}
model1 <- glm(depot ~ ., data = df1, family = 'binomial')

null_model1 <- glm(depot ~ 1, data = df1, family = 'binomial')

opt_model1 <- step(null_model1, direction = 'both',
                  scope = list(upper=model1, lower=null_model1), 
                  trace = FALSE)

```



### Check on quality: Model 1

```{r}
# Model coefficients
summary(opt_model1)
```


A better representation for the model results:

```{r}
tab_mod1 <-tbl_regression(opt_model1, exponentiate = TRUE) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p(t = 0.10) %>%
  bold_labels()
tab_mod1
```

```{r}
# Quality
check_model(opt_model1)
```



```{r, echo = FALSE, include=FALSE}
## Survival Analysis I
### Stepwise selection: `contrainte` "Oui" and "Non"
# Index column
df2$id <- 1:nrow(df2)

surv_obj <- Surv(df2$temps_depot, df2$depot)
fullmodel <- coxph(surv_obj ~ contrainte_interet + contrainte_internet +
         contrainte_ordi + contrainte_elect + contrainte_boulot + contrainte_fam +
         contrainte_sante + contrainte_note + contrainte_finance + contrainte_autr_etud +
         contrainte_sujet_complexe + contrainte_sujet_recherche + contrainte_donnees +
         contrainte_analyse_donnees + contrainte_encadrement + contrainte_emigration +
         contrainte_autre, id = id, data = df2)

cox_step <- stepAIC(fullmodel, direction = "both", trace = FALSE)
summary(cox_step)
```




```{r, echo = FALSE, include=FALSE}
## Survival Analysis II

### Stepwise selection: `contrainte` "Oui" only
# Index column
df3$id <- 1:nrow(df3)

surv_obj2 <- Surv(df3$temps_depot, df3$depot)
fullmodel2 <- coxph(surv_obj2 ~ contrainte_interet + contrainte_motivation + contrainte_internet +
         contrainte_ordi + contrainte_elect + contrainte_boulot + contrainte_fam +
         contrainte_sante + contrainte_note + contrainte_finance + contrainte_autr_etud +
         contrainte_sujet_complexe + contrainte_sujet_recherche + contrainte_donnees +
         contrainte_analyse_donnees + contrainte_encadrement + contrainte_emigration +
         contrainte_autre, id = id, data = df3)

cox_step2 <- stepAIC(fullmodel2, direction = "both", trace = FALSE)
summary(cox_step2)
```


# **Importing dataset 2 for Survival Analysis**

```{r}
df_mem2 <- read_excel('./Data_model22.xlsx')
```

# **Multivariate Analysis II**

```{r, echo=FALSE}
## **Data wrangling**

df_mem2$depot <- as.factor(df_mem2$depot)
df_mem2$recoded_temps_depot <- as.factor(df_mem2$recoded_temps_depot)
df_mem2$sexe <- as.factor(df_mem2$sexe)
df_mem2$loc_etud_sec <- as.factor(df_mem2$loc_etud_sec)
df_mem2$type_etab_term <- as.factor(df_mem2$type_etab_term)
df_mem2$compos_seuh <- as.factor(df_mem2$compos_seuh)
df_mem2$domaine_etud1 <- as.factor(df_mem2$domaine_etud1)
df_mem2$domaine_etud2 <- as.factor(df_mem2$domaine_etud2)
df_mem2$generation <- as.factor(df_mem2$generation)
df_mem2$autr_etud <- as.factor(df_mem2$autr_etud)
df_mem2$approche <- as.factor(df_mem2$approche)
df_mem2$qual_accomp <- as.factor(df_mem2$qual_accomp)
df_mem2$temps_emploi <- as.factor(df_mem2$temps_emploi)
df_mem2$contrainte <- as.factor(df_mem2$contrainte)

df_mem2$contrainte_interet <- as.integer(df_mem2$contrainte_interet)
df_mem2$contrainte_interet <- as.factor(df_mem2$contrainte_interet)

df_mem2$contrainte_motivation <- as.integer(df_mem2$contrainte_motivation)
df_mem2$contrainte_motivation <- as.factor(df_mem2$contrainte_motivation)

df_mem2$contrainte_ordi <- as.integer(df_mem2$contrainte_ordi)
df_mem2$contrainte_ordi <- as.factor(df_mem2$contrainte_ordi)

df_mem2$contrainte_elect <- as.integer(df_mem2$contrainte_elect)
df_mem2$contrainte_elect <- as.factor(df_mem2$contrainte_elect)

df_mem2$contrainte_internet <- as.integer(df_mem2$contrainte_internet)
df_mem2$contrainte_internet <- as.factor(df_mem2$contrainte_internet)

df_mem2$contrainte_boulot <- as.integer(df_mem2$contrainte_boulot)
df_mem2$contrainte_boulot <- as.factor(df_mem2$contrainte_boulot)

df_mem2$contrainte_fam <- as.integer(df_mem2$contrainte_fam)
df_mem2$contrainte_fam <- as.factor(df_mem2$contrainte_fam)

df_mem2$contrainte_sante <- as.integer(df_mem2$contrainte_sante)
df_mem2$contrainte_sante <- as.factor(df_mem2$contrainte_sante)

df_mem2$contrainte_note <- as.integer(df_mem2$contrainte_note)
df_mem2$contrainte_note <- as.factor(df_mem2$contrainte_note)

df_mem2$contrainte_finance <- as.integer(df_mem2$contrainte_finance)
df_mem2$contrainte_finance <- as.factor(df_mem2$contrainte_finance)

df_mem2$contrainte_autr_etud <- as.integer(df_mem2$contrainte_autr_etud)
df_mem2$contrainte_autr_etud <- as.factor(df_mem2$contrainte_autr_etud)

df_mem2$contrainte_sujet_recherche <- as.integer(df_mem2$contrainte_sujet_recherche)
df_mem2$contrainte_sujet_recherche <- as.factor(df_mem2$contrainte_sujet_recherche)

df_mem2$contrainte_sujet_complexe <- as.integer(df_mem2$contrainte_sujet_complexe)
df_mem2$contrainte_sujet_complexe <- as.factor(df_mem2$contrainte_sujet_complexe)

df_mem2$contrainte_donnees <- as.integer(df_mem2$contrainte_donnees)
df_mem2$contrainte_donnees <- as.factor(df_mem2$contrainte_donnees)

df_mem2$contrainte_analyse_donnees <- as.integer(df_mem2$contrainte_analyse_donnees)
df_mem2$contrainte_analyse_donnees <- as.factor(df_mem2$contrainte_analyse_donnees)

df_mem2$contrainte_encadrement <- as.integer(df_mem2$contrainte_encadrement)
df_mem2$contrainte_encadrement <- as.factor(df_mem2$contrainte_encadrement)

df_mem2$contrainte_emigration <- as.integer(df_mem2$contrainte_emigration)
df_mem2$contrainte_emigration <- as.factor(df_mem2$contrainte_emigration)

df_mem2$contrainte_autre <- as.integer(df_mem2$contrainte_autre)
df_mem2$contrainte_autre <- as.factor(df_mem2$contrainte_autre)

df_mem2$info_tfe <- as.factor(df_mem2$info_tfe)
df_mem2$difficulte_tfe <- as.factor(df_mem2$difficulte_tfe)
df_mem2$utilite_tfe <- as.factor(df_mem2$utilite_tfe)
df_mem2$projet_personnel <- as.factor(df_mem2$projet_personnel)
df_mem2$choix_etud <- as.factor(df_mem2$choix_etud)
df_mem2$education_fam <- as.factor(df_mem2$education_fam)

df_mem2$consommation_pps <- as.integer(df_mem2$consommation_pps)
df_mem2$consommation_pps <- as.factor(df_mem2$consommation_pps)

```



```{r, echo = FALSE, include=FALSE}
## Survival Analysis III

### Draw the subset for survival analysis III

df4 <- df_mem2 %>%
    select(colnames(df_mem),-var1, contrainte, -c('soutenance', 'temps_soutenance',
                                     'duree_etud_pot', 'duree_etud_reel',
                                     'duree_etud_diff', 'consent', 'preciser_etab_term',
                                     'etud_1', 'tranche_age', 'fac',
                                     'etud_2', 'comment_accomp',
                                     'preciser_autr_contr', 'domaine_etud2',
                                     'comment_difficulte', 'comment_utilite',
                                     'promotion', 'recoded_temps_depot')) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0))
  #filter(contrainte == "Oui")

```


```{r, echo = FALSE, include=FALSE}
### Stepwise selection: `contrainte` "Oui" and "Non"

# Index column
df4$id <- 1:nrow(df4)

surv_obj3 <- Surv(df4$temps_depot, df4$depot)
fullmodel3 <- coxph(surv_obj3 ~ contrainte_interet + contrainte_motivation + contrainte_internet +
         contrainte_ordi + contrainte_elect + contrainte_boulot + contrainte_fam +
         contrainte_sante + contrainte_note + contrainte_finance + contrainte_autr_etud +
         contrainte_sujet_complexe + contrainte_sujet_recherche + contrainte_donnees +
         contrainte_analyse_donnees + contrainte_encadrement + contrainte_emigration +
         contrainte_autre, id = id, data = df4)

cox_step <- stepAIC(fullmodel3, direction = "both", trace = FALSE)
summary(cox_step)
```


## Variable `temps_depot`: Normality test
```{r}
shapiro.test(df2$temps_depot) 

```

The variable `temps_depot` is not normally distributed as our p-value is inferior to 0.05.

## Survival Analysis IV

### Draw the subset for survival analysis IV

```{r}
df5 <- df_mem2 %>%
    select(colnames(df_mem),-var1, contrainte, -c('soutenance', 'temps_soutenance',
                                     'duree_etud_pot', 'duree_etud_reel',
                                     'duree_etud_diff', 'consent', 'preciser_etab_term',
                                     'etud_1', 'tranche_age', 'fac',
                                     'etud_2', 'comment_accomp',
                                     'preciser_autr_contr', 'domaine_etud2',
                                     'comment_difficulte', 'comment_utilite',
                                     'promotion', 'recoded_temps_depot')) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0)) %>%
  filter(contrainte == "Oui")

```



### Stepwise selection: `contrainte` "Oui" only

```{r}
df5$id <- 1:nrow(df5)

surv_obj4 <- Surv(df5$temps_depot, df5$depot)
fullmodel4 <- coxph(surv_obj4 ~ contrainte_interet + contrainte_motivation + contrainte_internet +
         contrainte_ordi + contrainte_elect + contrainte_boulot + contrainte_fam +
         contrainte_sante + contrainte_note + contrainte_finance + contrainte_autr_etud +
         contrainte_sujet_complexe + contrainte_sujet_recherche + contrainte_donnees +
         contrainte_analyse_donnees + contrainte_encadrement + contrainte_emigration +
         contrainte_autre, id = id, data = df5)

cox_step <- stepAIC(fullmodel4, direction = "both", trace = FALSE)
summary(cox_step)
```



### Graphical analysis - Kaplan-Meier Curve

```{r}
surv <- survfit(Surv(temps_depot, depot) ~ 1, data = df5)

plot(surv, col = c("red", "blue", "black", "green", "magenta"), lty = 1, xlab = "Time",
     ylab = "Survival Probability", main = "Kaplan-Meier Curve")
#abline(h=0.5, col="green")
```

```{r}
surv
```



### Checking the hypothesis of proportionality of the hazards
```{r log-log graphic}
suppressPackageStartupMessages(library(survminer))

ggcoxdiagnostics(cox_step, type = "martingale")
```

### Risk function graphic

```{r Risk function}
ggsurvplot(survfit(cox_step), data = df5, risk.table = TRUE, risk.table.col = "strata", pval = TRUE)
```



```{r, echo=FALSE}
## Another Survival Analysis _ General Model

df6 <- df_mem2 %>%
  select(-c("consent", "tranche_age", "preciser_etab_term", "fac", "etud_1", "domaine_etud2", "promotion", "recoded_temps_depot", "temps_soutenance", "duree_etud_reel", "etud_2", "soutenance", "duree_etud_pot", "duree_etud_diff", "comment_accomp", "preciser_autr_contr", "comment_difficulte", "comment_utilite")) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0)) %>%
  filter(contrainte == "Oui")

```



```{r, eval = FALSE, echo=FALSE, warning = FALSE}
### Stepwise selection: `contrainte` "Oui" only

df6$id <- 1:nrow(df6)

surv_obj5 <- Surv(df6$temps_depot, df6$depot)

fullmodel5 <- coxph(surv_obj5 ~ sexe + loc_etud_sec + type_etab_term + compos_seuh + domaine_etud1 + autr_etud + generation + approche + qual_accomp + temps_emploi + contrainte + info_tfe + difficulte_tfe + utilite_tfe + projet_personnel + choix_etud + consommation_pps + education_fam + contrainte_interet + contrainte_motivation + contrainte_internet +       contrainte_ordi + contrainte_elect + contrainte_boulot + contrainte_fam +
contrainte_sante + contrainte_note + contrainte_finance + contrainte_autr_etud + contrainte_sujet_complexe + contrainte_sujet_recherche + contrainte_donnees + contrainte_analyse_donnees + contrainte_encadrement + contrainte_emigration + contrainte_autre, id = id, data = df6)

cox_step3 <- stepAIC(fullmodel5, direction = "both", trace = FALSE)
summary(cox_step3)
```



## Survival analysis: sig. factors model LR + sig. contraintes

```{r}
df6$id <- 1:nrow(df6)

surv_obj5 <- Surv(df6$temps_depot, df6$depot)

fullmodel5 <- coxph(surv_obj5 ~ sexe + qual_accomp + contrainte + contrainte_motivation + contrainte_elect + contrainte_sante + contrainte_finance + contrainte_sujet_complexe, id = id, data = df6)

cox_step3 <- stepAIC(fullmodel5, direction = "both", trace = FALSE)
summary(cox_step3)
```

### Survival Analysis: factors LR

```{r, eval = FALSE, echo=FALSE, warning = FALSE}
### Stepwise selection: `contrainte` "Oui" only

df6$id <- 1:nrow(df6)

surv_obj5 <- Surv(df6$temps_depot, df6$depot)

fullmodel7 <- coxph(surv_obj5 ~ sexe + loc_etud_sec + type_etab_term + compos_seuh + domaine_etud1 + autr_etud + generation + approche + qual_accomp + temps_emploi + contrainte + info_tfe + difficulte_tfe + utilite_tfe + projet_personnel + choix_etud + consommation_pps + education_fam, id = id, data = df6)

cox_step4 <- stepAIC(fullmodel7, direction = "both", trace = FALSE)
summary(cox_step4)
```



