---
title: "Modélisation Mémoire - Régression Logistique et Analyse de survie"
author: "Davidson Adrien"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(scipen = 999)
```

# **Importing required packages**

```{r import packages}
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(officer))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(glmulti))
suppressPackageStartupMessages(library(flextable))
suppressPackageStartupMessages(library(performance))
suppressPackageStartupMessages(library(Rcpp))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(see))
suppressPackageStartupMessages(library(skimr))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(dlookr))
suppressPackageStartupMessages(library(report))
suppressPackageStartupMessages(library(ggstatsplot))
suppressPackageStartupMessages(library(gt))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(gtsummary))
suppressPackageStartupMessages(library(ggfortify))
suppressPackageStartupMessages(library(glmnet))
suppressPackageStartupMessages(library(survminer))

```

# **Importing the dataset for LR**

```{r import dataset}
df_mem <- read_excel('./Data_model2.xlsx')
df_mem2 <- read_excel('./Data_model22.xlsx')

```


# **Data wrangling**

## Adding full temps_depot column

```{r}
df_mem$temps_depot_complet <- df_mem2$temps_depot
```

## Factorizing the variables

```{r data wrangling: as.factor-ing the variables}
df_mem$depot <- as.factor(df_mem$depot)
df_mem$sexe <- as.factor(df_mem$sexe)
df_mem$loc_etud_sec <- as.factor(df_mem$loc_etud_sec)
df_mem$type_etab_term <- as.factor(df_mem$type_etab_term)
df_mem$compos_seuh <- as.factor(df_mem$compos_seuh)
df_mem$domaine_etud1 <- as.factor(df_mem$domaine_etud1)
df_mem$domaine_etud2 <- as.factor(df_mem$domaine_etud2)
df_mem$generation <- as.factor(df_mem$generation)
df_mem$autr_etud <- as.factor(df_mem$autr_etud)
df_mem$approche <- as.factor(df_mem$approche)
df_mem$qual_accomp <- as.factor(df_mem$qual_accomp)
df_mem$temps_emploi <- as.factor(df_mem$temps_emploi)
df_mem$contrainte <- as.factor(df_mem$contrainte)

df_mem$contrainte_interet <- as.integer(df_mem$contrainte_interet)
df_mem$contrainte_interet <- as.factor(df_mem$contrainte_interet)

df_mem$contrainte_motivation <- as.integer(df_mem$contrainte_motivation)
df_mem$contrainte_motivation <- as.factor(df_mem$contrainte_motivation)

df_mem$contrainte_ordi <- as.integer(df_mem$contrainte_ordi)
df_mem$contrainte_ordi <- as.factor(df_mem$contrainte_ordi)

df_mem$contrainte_elect <- as.integer(df_mem$contrainte_elect)
df_mem$contrainte_elect <- as.factor(df_mem$contrainte_elect)

df_mem$contrainte_internet <- as.integer(df_mem$contrainte_internet)
df_mem$contrainte_internet <- as.factor(df_mem$contrainte_internet)

df_mem$contrainte_boulot <- as.integer(df_mem$contrainte_boulot)
df_mem$contrainte_boulot <- as.factor(df_mem$contrainte_boulot)

df_mem$contrainte_fam <- as.integer(df_mem$contrainte_fam)
df_mem$contrainte_fam <- as.factor(df_mem$contrainte_fam)

df_mem$contrainte_sante <- as.integer(df_mem$contrainte_sante)
df_mem$contrainte_sante <- as.factor(df_mem$contrainte_sante)

df_mem$contrainte_note <- as.integer(df_mem$contrainte_note)
df_mem$contrainte_note <- as.factor(df_mem$contrainte_note)

df_mem$contrainte_finance <- as.integer(df_mem$contrainte_finance)
df_mem$contrainte_finance <- as.factor(df_mem$contrainte_finance)

df_mem$contrainte_autr_etud <- as.integer(df_mem$contrainte_autr_etud)
df_mem$contrainte_autr_etud <- as.factor(df_mem$contrainte_autr_etud)

df_mem$contrainte_sujet_recherche <- as.integer(df_mem$contrainte_sujet_recherche)
df_mem$contrainte_sujet_recherche <- as.factor(df_mem$contrainte_sujet_recherche)

df_mem$contrainte_sujet_complexe <- as.integer(df_mem$contrainte_sujet_complexe)
df_mem$contrainte_sujet_complexe <- as.factor(df_mem$contrainte_sujet_complexe)

df_mem$contrainte_donnees <- as.integer(df_mem$contrainte_donnees)
df_mem$contrainte_donnees <- as.factor(df_mem$contrainte_donnees)

df_mem$contrainte_analyse_donnees <- as.integer(df_mem$contrainte_analyse_donnees)
df_mem$contrainte_analyse_donnees <- as.factor(df_mem$contrainte_analyse_donnees)

df_mem$contrainte_encadrement <- as.integer(df_mem$contrainte_encadrement)
df_mem$contrainte_encadrement <- as.factor(df_mem$contrainte_encadrement)

df_mem$contrainte_emigration <- as.integer(df_mem$contrainte_emigration)
df_mem$contrainte_emigration <- as.factor(df_mem$contrainte_emigration)

df_mem$contrainte_autre <- as.integer(df_mem$contrainte_autre)
df_mem$contrainte_autre <- as.factor(df_mem$contrainte_autre)

df_mem$info_tfe <- as.factor(df_mem$info_tfe)
df_mem$difficulte_tfe <- as.factor(df_mem$difficulte_tfe)
df_mem$utilite_tfe <- as.factor(df_mem$utilite_tfe)
df_mem$projet_personnel <- as.factor(df_mem$projet_personnel)
df_mem$choix_etud <- as.factor(df_mem$choix_etud)
df_mem$education_fam <- as.factor(df_mem$education_fam)

df_mem$consommation_pps <- as.integer(df_mem$consommation_pps)
df_mem$consommation_pps <- as.factor(df_mem$consommation_pps)

```

```{r}
df_mem$temps_emploi <- relevel(df_mem$temps_emploi, ref = "Chomage")
df_mem$info_tfe <- relevel(df_mem$info_tfe, ref = "Pas_du_tout")
df_mem$utilite_tfe <- relevel(df_mem$utilite_tfe, ref = "NSP")
df_mem$choix_etud <- relevel(df_mem$choix_etud, ref = "Résigné")

```


## Partitions of the dataset for modelling purposes

```{r Partitions of the dataset for modelling purposes}
#Partitionning the dataset
var1 <- c('sexe', 'loc_etud_sec', 'type_etab_term', 'compos_seuh', 'domaine_etud1',
          'autr_etud', 'generation', 'approche', 'qual_accomp', 'temps_emploi',  
          'info_tfe', 'difficulte_tfe', 'utilite_tfe', 'projet_personnel',
          'choix_etud', 'consommation_pps', 'education_fam', 'contrainte_motivation', 'nombre_contrainte')

df1 <- df_mem %>%
  select(depot, var1)

```


# **Multivariate Analysis I**

## Logistic Regression

### LR: Depot vs Nombre_contrainte

```{r, echo=FALSE}
model_test <- glm(depot~nombre_contrainte, data = df1, family = 'binomial')
summary(model_test)
```

```{r, echo=FALSE}
tab_mod <-tbl_regression(model_test, exponentiate = TRUE) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p(t = 0.10) %>%
  bold_labels()
tab_mod
```


```{r exporting model results, include = FALSE}
sink("results_model_test.txt")
print(summary(model_test))
sink()

```

### Logistic regression full model

```{r}
model_full <- glm(depot~., data = df1, family = 'binomial')
summary(model_full)
```

```{r}
tab_mod_full <-tbl_regression(model_full, exponentiate = TRUE) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p(t = 0.10) %>%
  bold_labels()
tab_mod_full

```


```{r, include = FALSE}
sink("results_model_full2.txt")
print(tab_mod_full)
sink()
```

### Stepwise selection: logistic regression

```{r}
model_full <- glm(depot ~ ., data = df1, family = 'binomial')

null_model1 <- glm(depot ~ 1, data = df1, family = 'binomial')

model_opt <- step(null_model1, direction = 'both',
                  scope = list(upper=model_full, lower=null_model1), 
                  trace = FALSE)

```



### Check on quality: Model 1

```{r}
# Model coefficients
summary(model_opt)
```

### Test du rapport des maxima de vraisemblance

```{r}
lr_test <- anova(model_opt, test = 'LRT')
print(lr_test)
```

A better representation for the model results:

```{r}
tab_mod1 <-tbl_regression(model_opt, exponentiate = TRUE) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p(t = 0.10) %>%
  bold_labels()
tab_mod1
```


```{r, include = FALSE}
sink("results_model_opt.txt")
print(summary(model_opt))
sink()
```

```{r}
# Quality
check_model(model_opt)
```


# **Multivariate Analysis II**

# **Survival Analysis**


```{r, echo=FALSE}
## Another Survival Analysis _ General Model

df2 <- df_mem %>%
  select(-c("consent", "tranche_age", "preciser_etab_term", "fac", "etud_1", "domaine_etud2", "promotion", "temps_soutenance", "duree_etud_reel", "etud_2", "soutenance", "duree_etud_pot", "duree_etud_diff", "comment_accomp", "preciser_autr_contr", "comment_difficulte", "comment_utilite", "contrainte_interet", "contrainte_internet" , "contrainte_ordi" , "contrainte_elect" , "contrainte_boulot" , "contrainte_fam" , "contrainte_sante" , "contrainte_note" , "contrainte_finance" , "contrainte_autr_etud" , "contrainte_sujet_complexe" , "contrainte_sujet_recherche" , "contrainte_donnees" , "contrainte_analyse_donnees" , "contrainte_encadrement" , "contrainte_emigration" , "contrainte_autre", "contrainte")) %>%
  mutate(depot = ifelse(depot == "Oui", 1, 0))
  

df2$id <- 1:nrow(df2)
```

### **Time to submission factor distribution**

```{r, echo = FALSE}
hist(df_mem$temps_depot_complet, main = "Distribution du temps de soumission (ou non)",
     xlab = "Temps de soumission", ylab = "Fréquence", col = 'lightblue')
```

## **Kaplan-Meier Analysis Model specification**

### **Survival Analysis - Null Model**

```{r}
km_fit0 <- survfit(Surv(temps_depot_complet, depot) ~ 1, data = df2)
summary(km_fit0)
print(km_fit0)

```

```{r}
ggsurvplot(km_fit0, data = df2, risk.table=TRUE, conf.int=TRUE, ggtheme = theme_minimal())
```


### **Survival Analysis for Motivation factor**

```{r}
km_fit1 <- survfit(Surv(temps_depot_complet, depot) ~ contrainte_motivation, data = df2)
summary(km_fit1)
print(km_fit1)

```


```{r}
ggsurvplot(km_fit1, data = df2, risk.table=TRUE, conf.int=TRUE, pval = TRUE, pval.method = TRUE,
           ggtheme = theme_minimal())
```

```{r}
km_fit2 <- survfit(Surv(temps_depot_complet, depot) ~ sexe, data = df2)
summary(km_fit2)
print(km_fit2)

```


```{r}
ggsurvplot(km_fit2, data = df2, risk.table=TRUE, conf.int=TRUE, pval = TRUE, pval.method = TRUE,
           ggtheme = theme_minimal())
```


### **Cox Proportional Hazard Model Generation**

```{r, warning=FALSE}
surv_obj <- Surv(df2$temps_depot_complet, df2$depot)
fullmodel <- coxph(surv_obj ~ sexe + loc_etud_sec + type_etab_term + compos_seuh + domaine_etud1 + autr_etud + generation + approche + qual_accomp + temps_emploi + info_tfe + difficulte_tfe + utilite_tfe + projet_personnel + choix_etud + consommation_pps + education_fam + contrainte_motivation, id = id, data = df2)

cox_step <- stepAIC(fullmodel, direction = "both", trace = FALSE)
summary(cox_step)
```

### Significant variables
```{r}
significant_vars <- summary(cox_step)$coefficients[summary(cox_step)$coefficients[, "Pr(>|z|)"] < 0.1, ]
significant_vars
```

```{r}
tab_mod_surv <-tbl_regression(cox_step, exponentiate = TRUE) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p(t = 0.10) %>%
  bold_labels()
tab_mod_surv
```

